A Combinatorial Search of Parameterized Quantum Circuit Learning for Chemical Applications

### 2024-12-02 manuscript/main.tex
Manuscript + Conversion to Qiskit progress

\newpage
\section{Introduction}
\cite{prasad_applications_2024}

A1, A2, M-Encoding \cite{suzuki_predicting_2020}
M-Encoding \cite{mitarai_quantum_2018}

Ansaetze \cite{sim_expressibility_2019}
``In particular, a substantial improvement in performance of two-qubit gates in a ring or all-to-all connected arrangement, compared to thatof those on a line, is observed.''

IQP \cite{bremner_average-case_2015}

\cite{hatakeyama-sato_quantum_2023}
\cite{krenn_artificial_2023}
\cite{benedetti_parameterized_2019}
\cite{biamonte_quantum_2017}

As current quantum hardware grows, much attention is brought to find applicable uses for the potentially powerful new technology while it continues to grow. Current research in the areas of developing quantum algorithms for use with current noisy intermediate-scale quantum (NISQ) hardware lays the foundation for understanding and applying quantum computing. These near-term algorithms have seen a wide deployment across many areas of science and technology as the field grows. Computational chemistry is a natural such candidate for quantum computing applications due to the quantum nature of many problems in the field. An especially important problem is that of calculating the bond separation energies of (molecules?). Efficient and accurate predictions of these values can aid research in material science and drug discovery.\par

One of the most promising near-term applications of quantum computing within any field has been quantum machine learning (QML). The advantages of QML comes largely from its ability to be readily deployed on quantum computers with few qubits, and its potential to speedup problems with minimal loss of accuracy. The number of qubits needed for a QML model is determined primarily by the number of features in the dataset. As there exist data science techniques developed in the realm of classical machine learning to reduce feature spaces, they can be adopted over to reduce problems to a suitable number of qubits. Further, since QML only needs to train once, it has potential to perform faster when compared to other current NISQ algorithms such as the variational quantum eigensolver, which needs to be trained separately for each problem. \par

Quantum machine learning has been widely applied to classification tasks with much success. However, little research has been done to understand its capabilities for regression. Previous works have set important groundwork to the understanding of QML for regression by theoretically formulating upper bounds for the generalization error. Most notably, this research suggests that with an efficiently implemented QML model can achieve good generalizability with modest training data sizes.  This highlights a potential for QML models to outperform and out-generalize classical models. Further understanding and verification of this, however, requires numerical validation and demonstration, and while some experimental work has been done, it remains limited to few qubits with little practical application.\par

Our work expands on the current knowledge base by providing a comprehensive exploration of QML for regression. Specfically, we will by applying QML to the problem of calculating bond dissociation enthalpies of various molecules. This demonstrates 
a practical application of regression in QML with many applications. The efficient and accurate prediction of these chemical properties can aid research in material science and drug discovery with far reaching implications. Our work serves to benchmark QML and test its capabilities within this context. \par

\section{Methods}
Quantum machine learning is a hybrid classical-quantum approach utilizing parameterized quantum circuits. Typical hybrid algorithms consist of shallow circuit depth of one qubit rotations and two qubit entangling gates. In the context of machine learning, we can subdivide our circuit into three components: the encoder, the variational circuit, and the act of measuring to estimate predictions. The encoder takes our $n$-dimensional data $\bm x \in \R^n$ and encodes it into a unitary feature map $U(\bm x)$ in our working Hilbert Space. This unitary acts on our initialized state $\ket{0}^{\otimes n'}$ to produce the state $U(\bm x) \ket{0}^{\otimes n'}$. It is usual to take $n' = n$ and for each feature $x_i$ to correspond to a qubit. With $n < n'$, features can be repeated across the qubits present in the system. Further elaboration will be discussed with specific encoding circuits. After the encoder produces the state $U(\bm x)\ket{0}^{\otimes n}$, we then tune the state with a parameterized unitary $U(\bm \theta)$ to transform the state into a meaningful prediction. Finally, some configuration of qubits are measured to produce expectation values $\hat y_k$ for our data labels $y_i$ which we then use to minimize a loss function. An obstacle in creating quantum models is there is no way to determine the best encoding or variational circuit for the problem. As such, this paper explores many different encoding and variational circuits. \par

@@

\section{Results}
\subsection{Model Calibration Using Function fitting} 
We used the function fitting to determine if our models were implemented correctly.

\subsubsection{All (5 + 16 qubit)}
Figure \ref{fig:5qubit_Func_funcfit_R2} shows every encoder and ansatz pair explored for the function fitting datasest with 5 qubits.
\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{\textwidth}
		\centering
		\includegraphics[width=\textwidth]{images/Function_Fitting/5qubit_Linear_funcfit_R2.png}
		\caption{}
		\label{fig:5qubit_Linear_funcfit_R2}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{\textwidth}
		\centering
		\includegraphics[width=\textwidth]{images/Function_Fitting/5qubit_Quadratic_funcfit_R2.png}
		\caption{}
		\label{fig:5qubit_Quadratic_funcfit_R2}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{\textwidth}
		\centering
		\includegraphics[width=\textwidth]{images/Function_Fitting/5qubit_Sine_funcfit_R2.png}
		\caption{}
		\label{fig:5qubit_Sine_funcfit_R2}
	\end{subfigure}
	\caption{}
	\label{fig:5qubit_Func_funcfit_R2}
\end{figure}

Figure \ref{fig:16qubit_Func_funcfit_R2} shows every encoder and ansatz pair explored for the function fitting datasest with 16 qubits.

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{\textwidth}
		\centering
		\includegraphics[width=\textwidth]{images/Function_Fitting/16qubit_Linear_funcfit_R2.png}
		\caption{}
		\label{fig:16qubit_Linear_funcfit_R2}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{\textwidth}
		\centering
		\includegraphics[width=\textwidth]{images/Function_Fitting/16qubit_Quadratic_funcfit_R2.png}
		\caption{}
		\label{fig:16qubit_Quadratic_funcfit_R2}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{\textwidth}
		\centering
		\includegraphics[width=\textwidth]{images/Function_Fitting/16qubit_Sine_funcfit_R2.png}
		\caption{}
		\label{fig:16qubit_Sine_funcfit_R2}
	\end{subfigure}
	\caption{}
	\label{fig:16qubit_Func_funcfit_R2}
\end{figure}

\subsubsection{RUD + AL (5 + 16 qubit)}\label{subsection:func_depth}
The best, median, and worst 5 were chosen from the 16 qubit data.

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{\textwidth}
		\centering
		\includegraphics[width=\textwidth]{images/Function_Fitting/16qubit_Linear_RUD_AL.png}
		\caption{}
		\label{fig:16qubit_Linear_RUD_AL}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{\textwidth}
		\centering
		\includegraphics[width=\textwidth]{images/Function_Fitting/16qubit_Quadratic_RUD_AL.png}
		\caption{}
		\label{fig:16qubit_Quadratic_RUD_AL}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{\textwidth}
		\centering
		\includegraphics[width=\textwidth]{images/Function_Fitting/16qubit_Sine_RUD_AL.png}
		\caption{}
		\label{fig:16qubit_Sine_RUD_AL}
	\end{subfigure}
	\caption{}
	\label{fig:16qubit_Func_RUD_AL}
\end{figure}

\subsubsection{Learning curve}
The best models from Section \ref{subsection:func_depth} were then used to compare them with classical machine learning models using learning curves, where the test set was held to 10 points, while the training data was varied from 10-80 points.

For the linear model with 16 qubits, the IQP encoding with the Full Pauli CRZ ansatz layer outperforms the classical models at 10\% of the training data, as show in  Fig. \ref{fig:linear_learning_curves}.
The quantum model is less overfit and has a more accurate test score than any of the classical models. 
As the training set size increases, the quantum model maintains an average train and test R$^{2}$ of 0.98, while the classical models maintain an average train and test R$^{2}$ of 0.99.
\begin{figure}[H]
	\centering
	\includegraphics[width=\textwidth]{images/Function_Fitting/linear_learning_curves.png}
	\caption{Linear learning curve}
	\label{fig:linear_learning_curves}
\end{figure}

A2 HWE-CZ
Performs better than all but 
\begin{figure}[H]
	\centering
	\includegraphics[width=\textwidth]{images/Function_Fitting/sine_learning_curves.png}
	\caption{sine learning curve}
	\label{fig:sine_learning_curves}
\end{figure}

\subsubsection{Error mitigation}
FakeQuebec backend
ZNE linear and Richardson
TREX
Without

\subsubsection{Real device}
IBM Quebec!

\subsection{BSE}

\subsubsection{Classical testing}
After an extensive search of the possible moleuclar representations for the BSE dataset, we decided to use Morgan fingerprints.
To map this set of features down to 5 and 16 qubits, we analyzed using SHapley Additive exPlanation (SHAP) values and principal component analysis.
Based of of the analysis from Fig. \ref{fig:Feat_redR2}, we chose to use the set of features reduced using PCA due to the more accurate, albiet overfit data.
\begin{figure}[H]
	\centering
	\includegraphics[width=\textwidth]{images/BSE/Feat_redR2.png}
	\caption{BSE Feature Reduction}
	\label{fig:Feat_redR2}
\end{figure}
\subsubsection{All (5 + 16 qubit)}
Using the aforementioned datasets, along with the 168 encoder-ansatz pairs we analyzed the cost of modeling the BSE dataset using the 5 qubit feature set first.

\begin{figure}[H]
	\centering
	\includegraphics[width=\textwidth]{images/BSE/5qubit_BSE_MAE.png}
	\caption{5 qubit BSE MAE (kcal/mol)}
	\label{fig:5qubit_BSE_MAE}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=\textwidth]{images/BSE/16qubit_BSE_MAE.png}
	\caption{16 qubit BSE MAE (kcal/mol)}
	\label{fig:16qubit_BSE_MAE}
\end{figure}

Using this data, we analyzed the mean time per training iteration versus the MAE and circuit depth. 
We found that within the 168 encoder-ansatz pairs, there existed a bimodal distribution, where there were a set of 70 circuits with the mean time per iteration greater than 100 seconds and those with less.
Going forward, we will only use the set of circuits that belong to the set with mean iteration times less than 100 seconds due to cost considerations.
\begin{figure}[H]
	\centering
	\includegraphics[width=\textwidth]{images/BSE/AllBSE5_circuitdepth_MAE_vs_iterationtime.png}
	\caption{5 qubit BSE MAE (kcal/mol)}
	\label{fig:AllBSE5_circuitdepth_MAE_vs_iterationtime}
\end{figure}

For the 70 circuits that belong to those with iterations less than 100 seconds for the BSE 5 qubit set, we used these for the 16 qubit as well and found that...

\subsubsection{RUD + AL (5 + 16 qubit)}
Based off of the improvements shown in the function fitting dataset, we only used the best encoding ansatz pair for the real datasets
For the 5 qubit, it is  M-M-CZ HWE-CNOT

\begin{figure}[H]
	\centering
	\includegraphics[width=\textwidth]{images/BSE/BSE5_RUDAL.png}
	\caption{5 qubit BSE MAE (kcal/mol) AL RUD}
	\label{fig:	BSE5_RUDAL}
\end{figure}

For 16 qubit A2 Hadamard

\subsubsection{Error mitigation}
\subsubsection{Real device}

\subsection{DDCC}

\subsubsection{Classical testing}
\subsubsection{All (5 + 16 qubit)}
\subsubsection{RUD + AL (5 + 16 qubit)}
\subsubsection{Error mitigation}
\subsubsection{Real device}

>>>>

\newpage
\section{Introduction}
[quantum machine learning]

[parameterized circuits for chemical applications]

[What we introduce]

Our work includes the single and double encoding layers found in the paper of Suzuki and Katouda\cite{suzuki_predicting_2020}, including the encoding of Mitarai \textit{et al.}\cite{mitarai_quantum_2018}, along with using instantanious quantum polynomial (IQP) as an encoding layer. \cite{bremner_average-case_2015}
All of our variational (ansatz) layers can be found in the study of Sim \textit{et al.}\cite{sim_expressibility_2019}, where the expressiblity and entanglement capability of these circuits is analyzed.

\cite{krenn_artificial_2023}

QML\cite{biamonte_quantum_2017}:
\begin{itemize}
	\item ``Quantum machine learning software makes use of quantum algorithms as part of a larger implementation. By analysing the steps that quantum algorithms prescribe, it becomes clear that they have the potential to outperform classical algorithms for specific problems (that is, reduce the number of steps required). This potential is known as quantum speedup.''
	\item ``The notion of a quantum speedup depends on whether one takes a formal computer science perspective—which demands mathematical proofs—or a perspective based on what can be done with realistic, finitesize devices—which requires solid statistical evidence of a scaling advantage over some finite range of problem sizes. For the case of quantum machine learning, the best possible performance of classical algorithms is not always known.''
	\item DO WE SEE THIS: ``Determination of a scaling advantage contrasting quantum and classical machine learning would rely on the existence of a quantum computer and is called a ‘benchmarking’ problem. Such advantages could include improved classification accuracy and sampling of classically inaccessible systems. Accordingly, quantum speedups in machine learning are currently characterized using idealized measures from complexity theory: query complexity and gate complexity (see Box 1 and Box 1 Table). Query complexity measures the number of queries to the information source for the classical or quantum algorithm. A quantum speedup results if the number of queries needed to solve a problem is lower for the quantum algorithm than for the classical algorithm. To determine the gate complexity, the number of elementary quantum operations (or gates) required to obtain the desired result are counted.''
	\item Input problem: ``Classical data must be input before being processed on a quantum computer. This ‘input problem’ often has little overhead but can present a serious bottleneck for certain algorithms. Likewise, the ‘output problem’ is faced when reading out data after being processed on a quantum device. Like the input problem, the output problem often causes a noticeable operational slowdown.''
	\item  QML challenges: ``These hardware challenges are technical in nature, and clear paths exist towards overcoming them. They must be overcome, however, if quantum machine learning is to become a ‘killer app’ for quantum computers. As noted previously, most of the quantum algorithms that have been identified face a number of caveats that limits their applicability. We can distill the caveats mentioned above into four fundamental problems. (1) The input problem. Although quantum algorithms can provide dramatic speedups for processing data, they seldom provide advantages in reading data. This means that the cost of reading in the input can in some cases dominate the cost of quantum algorithms. Understanding this factor is an ongoing challenge. (2) The output problem. Obtaining the full solution from some quantum algorithms as a string of bits requires learning an exponential number of bits. This makes some applications of quantum machine learning algorithms infeasible. This problem can potentially be sidestepped by learning only summary statistics for the solution state. (3) The costing problem. Closely related to the input/outputproblems, at present very little is known about the true number of gates required by quantum machine learning algorithms. Bounds on the complexity suggest that for sufficiently large problems they will offer huge advantages, but it is still unclear when that crossover point occurs. (4) The benchmarking problem. It is often difficult to assert that a quantum algorithm is ever better than all known classical machine algorithms in practice because this would require extensive benchmarking against modern heuristic methods. Establishing lower bounds for quantum machine learning would partially address this issue.''
\end{itemize}

PQCs\cite{benedetti_parameterized_2019}:
\begin{itemize}
	\item ``Parameterized quantum circuits (PQCs) offer a concrete way to implement algorithms and demonstrate quantum supremacy in the NISQ era. PQCs are typically composed of fixed gates, e.g. controlled NOTs, and adjustable gates, e.g. qubit rotations. Even at low circuit depth, some classes of PQCs are capable of generating highly non-trivial outputs. For example, under well-believed complexity-theoretic assumptions, the class of PQCs called instantaneous quantum polynomial-time cannot be efficiently simulated by classical resources (see Lund et al [3] and Harrow and Montanaro [4] for accessible Reviews of quantum supremacy proposals). The demonstration of quantum supremacy is an important milestone in the development of quantum computers. In practice, however, it is highly desirable to demonstrate a quantum advantage on applications. The main approach taken by the community consists in formalizing problems of interest as variational optimization problems and use hybrid systems of quantum and classical hardware to find approximate solutions. The intuition is that by implementing some subroutines on classical hardware, the requirement of quantum resources is significantly reduced, particularly the number of qubits, circuit depth, and coherence time. Therefore, in the hybrid algorithmic approach NISQ hardware focuses entirely on the classically intractable part of the problem.''
\end{itemize}

PQCs \cite{suzuki_predicting_2020}:
\begin{itemize}
	\item ``More recently, quantum machine learning (QML) [26–30] is a rapidly growing research field that combines near-term quantum algorithms and machine learning techniques. In particular, parameterized quantum circuits (PQCs) have been considered as machine learning models with high expressive power within the hybrid quantum–classical framework [31, 32]. PQCs are typically composed of fixed quantum gates (e.g., qubit rotations and entangling gates) in a shallow circuit layout, with variable parameters optimized in a classical feedback loop.''
	\item ``To our knowledge, however, the application of QML to regression tasks has not been fully investigated in the literature. It remains unclear what kinds of quantum states should be used in order to generate the feature map with high expressibility that is suited for real-world data sets.''
	\item ``quantitative structure–toxicity relationship (QSTR) models for predicting the toxicity of 221 phenols. While there are a variety of QSAR/QSTR models (e.g., 3D-QSAR [1, 4]),asafirst step we employ QSAR/QSTR models including molecular descriptors such as hydrophobicity, acidity constant, and frontier orbital energies. There have been quantum computations in biochemical and pharmaceutical areas, such as protein folding [43–45], molecular similarity [46], and biological data [47]; yet, there has been no study on quantum application to QSAR modeling, albeit an important part of ligand-based computer-aided drug design.''
	\item ``PQCs have been regarded as machine learning models with high expressive power within the framework of the hybrid quantum–classical approach. PQCs are usually composed of one-qubit rotations and two-qubit entangling operations in a shallow circuit layout, with parameters optimized in a feedback loop.''
	\item ``Combining near-term quantum algorithms and machine learning, QML using the framework of PQCs is sometimes referred to as quantum circuit learning (QCL)''
	\item ``PQCs consist of three components: the encoder circuit, the variational circuit, and the measurement for the estimation of the loss function''

\end{itemize}

PQCs \cite{hatakeyama-sato_quantum_2023}:
\begin{itemize}
	\item What they do: ``ntroduce quantum circuit learning (QCL) as an emerging regression algorithm for chemo- and materials-informatics. The supervised model, functioning on the rule of quantum mechanics, can process linear and smooth non-linear functions from small datasets (<100 records). Compared with conventional algorithms, such as random forest, support vector machine, and linear regressions, the QCL can offer better predictions with some one-dimensional functions and experimental chemical databases''
	\item ``Here, we introduce quantum circuit learning (QCL), an emerging algorithm for supervised learning.15–18 QCL works on the rule of quantum mechanics. It can predict various parameters likewise to classical models. The current quantum systems (noisy intermediate-scale quantum computers: NISQ)19,20 face the problems of calculation noise and the limited processable number of information units (qubits).''
	\item ``Since QCL is a frontier for machine learning, few reports have been published on authentic regression tasks. The success of prediction with a toxicity dataset of organic molecules was reported,17 whereas the study was still conceptual. The prediction processes and advantages have been unclear, especially from chemical and material viewpoints. Here, we conducted a more comprehensive study on QCL with standard datasets of one-dimensional functions and chemical properties. Both simulated and actual quantum computing was undertaken to clarify the challenges of QCL. Various hyperparameters were optimized for higher accuracy. The comparison with conventional models contrasted the benefits of the new approach: capable of learning both linear and non-linear functions even from small datasets. The property was also favorable for predicting the so-called extrapolating data region, which is essential for chemical and material research.''
	\item They cite \cite{takaki_learning_2021,suzuki_predicting_2020,mitarai_quantum_2018} as examples of PQCs/quantum circuit learning for regression task
	\item Justification for starting with 1D regression: ``Preliminary optimization of quantum circuits15,17 and our optimization revealed that the following configuration was concise and practical for regression tasks (for details, see results and explanations in Fig. S1''
\end{itemize}

\section{Methods}
Use a mix of  \cite{hatakeyama-sato_quantum_2023} and \cite{suzuki_predicting_2020} for the methods.

Format from \cite{suzuki_predicting_2020}:
\begin{itemize}
	\item Feature set $\mathbf{x}=(x_{1}, x_{2}, \ldots, x_{d})^{T} \in \mathbb{R}^{d}$
	\item Feature map (encoding) $U_{\Phi(\mathbf{x})}$ produces the quantum state $U_{\Phi(\mathbf{x})}\ket{0}^{\otimes n}$
	\item Variational circuit (ansatz) $U(\mathbf{\theta})$, leads to the parameterized quantum circuit $U(\mathbf{\theta})U_{\Phi(\mathbf{x})}\ket{0}^{\otimes n}$
\end{itemize}

Quantum machine learning is a hybrid classical-quantum approach utilizing parameterized quantum circuits. Typical hybrid algorithms consist of shallow circuit depth of one qubit rotations and two qubit entangling gates. In the context of machine learning, we can subdivide our circuit into three components: the encoder, the variational circuit, and the act of measuring to estimate predictions. The encoder takes our $n$-dimensional data $\bm x \in \R^n$ and encodes it into a unitary feature map $U(\bm x)$ in our working Hilbert Space. This unitary acts on our initialized state $\ket{0}^{\otimes n'}$ to produce the state $U(\bm x) \ket{0}^{\otimes n'}$. It is usual to take $n' = n$ and for each feature $x_i$ to correspond to a qubit. With $n < n'$, features can be repeated across the qubits present in the system. Further elaboration will be discussed with specific encoding circuits. After the encoder produces the state $U(\bm x)\ket{0}^{\otimes n}$, we then tune the state with a parameterized unitary $U(\bm \theta)$ to transform the state into a meaningful prediction. Finally, some configuration of qubits are measured to produce expectation values $\hat y_k$ for our data labels $y_i$ which we then use to minimize a loss function. An obstacle in creating quantum models is there is no way to determine the best encoding or variational circuit for the problem. As such, this paper explores many different encoding and variational circuits. \par

@@

\section{Results}

Ansaetze analysis \cite{sim_expressibility_2019}
``In particular, a substantial improvement in performance of two-qubit gates in a ring or all-to-all connected arrangement, compared to thatof those on a line, is observed.''
